---
title: "Chapter 6"
author: "Robert W. Walker"
date: "2022-09-26"
categories: [R]
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, comment=NA, prompt=FALSE, fig.height=6, fig.width=6.5, fig.retina = 3, dev = 'svg', dev.args = list(bg = "transparent"))
options(scipen=7)
```

## 6.5.2 Data exercises

Use the same `health_insurance` data set from this chapter to answer these questions.

```{r}
url <- "http://peopleanalytics-regression-book.org/data/health_insurance.csv"
health_insurance <- read.csv(url)
```


1. Complete the full stratified approach to modeling the three product choices that was started in Section 6.2. Calculate the coefficients, odds ratios and p-values in each case.

```{r}
library(fastDummies)
health_insurance <- fastDummies::dummy_cols(health_insurance, "product")
B_model <- glm(
  formula = product_B ~ age + gender + household + 
    position_level + absent, 
  data = health_insurance, 
  family = "binomial"
)
# summary
summary(B_model)
exp(summary(B_model)$coefficients[,1])
C_model <- glm(
  formula = product_C ~ age + gender + household + 
    position_level + absent, 
  data = health_insurance, 
  family = "binomial"
)
# summary
summary(C_model)
exp(summary(C_model)$coefficients[,1])
```


2. Carefully write down your interpretation of the odds ratios from the previous question.

*The relative odds of B increase by 6.4% for each additional year of age.  The relative odds of B decrease by almost 90% for males.  The relative odds of B increase by 110% for Non-binary subjects.  The relative odds of B decrease by 63% for each member of a household.  The relative odds of B decrease by 25% for each position level and the relative odds of B increase by just under one percent for each day absent.*

*The relative odds of C increase by 11.8% for each additional year of age.  The relative odds of C increase by over 200% for males.  The relative odds of C decrease by 19% for Non-binary subjects.  The relative odds of C increase by 79% for each member of a household.  The relative odds of C decrease by 4% for each position level and the relative odds of C decrease by 0.006 percent for each day absent.*

3. Run a multinomial logistic regression model on the product outcome using Product B as reference. Calculate the coefficients, ratios and p-values in each case.

```{r}
library(nnet)
health_insurance$product <- relevel(as.factor(health_insurance$product), ref="B")
mod.MNL <- nnet::multinom(product ~ age+household+position_level+gender+absent, data=health_insurance)
summary(mod.MNL)
# calculate z-statistics of coefficients
z_stats <- summary(mod.MNL)$coefficients/
  summary(mod.MNL)$standard.errors
z_stats
# convert to p-values
p_values <- (1 - pnorm(abs(z_stats)))*2
# display p-values in transposed data frame
data.frame(t(p_values))
odds_ratios <- exp(summary(mod.MNL)$coefficients)
data.frame(t(odds_ratios))
```


4. Verify that the coefficients for Product C against reference Product B matches those calculated in Section 6.3.3.

*They do.*

5. Carefully write down your interpretation of the odds ratios calculated in the previous question.

*For each unit of age, the relative odds of A decrease by 22% and the relative odds of C increase by 2.6%.  For each member of the household, the relative odds of A increase by 163% and the relative odds of C increase by 223%.  For each unit of position level, the relative odds of A increase by 51.5% and increase for C by 22.4%.  Males have relative odds 983% higher for A and 1093% higher for C.  Non-binary has 22.4% lower relative odds for A and 78.2% lower relative odds for C.  Each day of absence decreases relative odds of A by 1.2% and C by .84%.*

6. Use the process described in Section 6.4.1 to simplify the multinomial model in Question 3.

```{r}
mod.MNL <- nnet::multinom(product ~ age+household+position_level+gender, data=health_insurance)
summary(mod.MNL)
```

*Dropping `absent` is justified following this procedure as the text demonstrates.  Next would be position level or gender and neither are justified.*
